Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-04-02 manuscript/manuscript.Rmd
WIP on central normality tests.

```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "CHECK. ")

residual <- residual_error <- threshold <- NULL

if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))){

  set.seed(95)

  # Generate table of asymmetric generalised normal distribution parameters.
  n_distributions <- 1000

  # Add outliers, between 0.00 and 0.10.
  n_outliers <- 10
  outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)

  x <- mapply(
    function(x, n, alpha, beta, outlier_fraction){

      parameters <- list("n" = n, "alpha"=alpha, "beta" = beta)

            # Generate outlier values that are smaller than Q1 - 1.5 IQR or larger
            # than Q3 + 1.5 IQR.
            x_random <- stats::runif(n_draw, min=-2.0, max=2.0)
            outlier <- numeric(n_draw)
            if(any(x_random < 0)){
              outlier[x_random < 0] <- q_lower - 1.5 * interquartile_range + x_random[x_random < 0] * interquartile_range
            }

            if(any(x_random >= 0)){
              outlier[x_random >= 0] <- q_upper + 1.5 * interquartile_range + x_random[x_random >= 0] * interquartile_range
            }

      return(x)
    },
    x = x,
    n = n,
    alpha = alpha,
    beta = beta,
    MoreArgs=list("outlier_fraction"=outlier_fraction),
    SIMPLIFY=FALSE,
    USE.NAMES=FALSE)

  x <- unlist(x, recursive = FALSE)
  x <- mapply(
    function(ii, x){
      x$distribution_id <- ii

      return(x)},
    ii = seq_along(x),
    x = x,
    SIMPLIFY = FALSE)

  compute_residuals <- function(
    x,
    n_sample = 200){

    # Set interpolation points.
    p_sample <- (seq_len(n_sample) - 1/3) / (n_sample + 1/3)

    # Determine residuals for Box-Cox transformations.
    transformer_bc <- suppressWarnings(power.transform::find_transformation_parameters(
      x = x$x,
      method = "box_cox",
      shift = TRUE,
      robust = TRUE,
      weight_method = "original_cosine",
      k1 = 0.76,
      k2 = 0.95))

    residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_bc)

    residual <- stats::approx(
      x = residual_data$p,
      y = abs(residual_data$residual),
      xout = p_sample,
      rule = 2,
      ties = max)$y

    residual_bc <- data.table::data.table(
      "distribution_id" = x$distribution_id,
      "p" = p_sample,
      "residual" = residual,
      "method" = "box_cox",
      "has_outliers" = x$parameters$k > 0)

    # Determine residuals for Yeo-Johnson transformations.
    transformer_yj <- suppressWarnings(power.transform::find_transformation_parameters(
      x = x$x,
      method = "yeo_johnson",
      shift = TRUE,
      robust = TRUE,
      weight_method = "original_cosine",
      k1 = 0.89,
      k2 = 0.93))

    residual_data <- power.transform::get_residuals(x = x$x, transformer = transformer_yj)

    residual <- stats::approx(
      x = residual_data$p,
      y = abs(residual_data$residual),
      xout = p_sample,
      rule = 2,
      ties = max)$y

    residual_yj <- data.table::data.table(
      "distribution_id" = x$distribution_id,
      "p" = p_sample,
      "residual" = residual,
      "method" = "yeo_johnson",
      "has_outliers" = x$parameters$k > 0)

    return(data.table::rbindlist(list(residual_bc, residual_yj)))
  }

  # Compute all data in parallel.
  data <- parallel::parLapply(
    cl=cl,
    X=x,
    fun=compute_residuals)

  # data <- lapply(
  #   x,
  #   compute_residuals)

  data <- data.table::rbindlist(data)

  saveRDS(
    object=data,
    file=file.path(manuscript_dir, "residual_plot.RDS"))

} else {
  data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))
}

data$method <- factor(
  x = data$method,
  levels = c("box_cox", "yeo_johnson"),
  labels = c("Box-Cox", "Yeo-Johnson"))

outlier_free_data <- data[has_outliers == FALSE, list("residual_error"=stats::quantile(residual, probs=0.99)), by=c("p", "method")]

data <- data[, list(
  "residual_50"=stats::quantile(residual, probs=0.50),
  "residual_90"=stats::quantile(residual, probs=0.90),
  "residual_95"=stats::quantile(residual, probs=0.95),
  "residual_99"=stats::quantile(residual, probs=0.99),
  "residual_max"=max(residual)),
  by=c("p", "method")]

data <- data.table::melt(
  data = data,
  id.vars = c("p", "method"),
  variable.name = "threshold",
  value.name = "residual_error")

data$threshold <- factor(
  x = data$threshold,
  levels = c("residual_50", "residual_90", "residual_95", "residual_99", "residual_max"),
  labels = c("50 %", "90 %", "95 %", "99 %", "100 %"))

p_bc <- ggplot2::ggplot(
  data = data[method == "Box-Cox"],
  mapping = ggplot2::aes(
    x = p,
    y = residual_error,
    colour = threshold))
p_bc <- p_bc + plot_theme
p_bc <- p_bc + ggplot2::geom_line()
p_bc <- p_bc + ggplot2::geom_line(
  data = outlier_free_data[method == "Box-Cox"],
  mapping = ggplot2::aes(
    x = p,
    y = residual_error),
  colour = "gray40",
  linetype = "dashed")
p_bc <- p_bc + ggplot2::scale_colour_discrete(
  name = "Box-Cox threshold",
  type=c("50 %" = "#ABC6E2", "90 %" = "#779EC6", "95 %" = "#4E79A7", "99 %" = "#346394", "100 %" = "#1D4D7E"))
p_bc <- p_bc + ggplot2::xlab("empirical probability")
p_bc <- p_bc + ggplot2::ylab("absolute residual error")
p_bc <- p_bc + ggplot2::coord_cartesian(
  xlim = c(0, 1),
  ylim = c(0, 1))
p_bc <- p_bc + ggplot2::theme(
  axis.title.x = ggplot2::element_blank(),
  axis.text.x = ggplot2::element_blank())

p_yj <- ggplot2::ggplot(
  data = data[method == "Yeo-Johnson"],
  mapping = ggplot2::aes(
    x = p,
    y = residual_error,
    colour = threshold))
p_yj <- p_yj + plot_theme
p_yj <- p_yj + ggplot2::geom_line()
p_yj <- p_yj + ggplot2::geom_line(
  data = outlier_free_data[method == "Yeo-Johnson"],
  mapping = ggplot2::aes(
    x = p,
    y = residual_error),
  colour = "gray40",
  linetype = "dashed")
p_yj <- p_yj + ggplot2::scale_colour_discrete(
  name = "Yeo-Johnson threshold",
  type=c("50 %" = "#FFBD7D", "90 %" = "#FFA954", "95 %" = "#F28E2B", "99 %" = "#CD6B0B", "100 %" = "#A25000"))
p_yj <- p_yj + ggplot2::xlab("empirical probability")
p_yj <- p_yj + ggplot2::ylab("absolute residual error")
p_yj <- p_yj + ggplot2::coord_cartesian(
  xlim = c(0, 1),
  ylim = c(0, 1))

p <- egg::ggarrange(
  plots=list(
    p_bc,
    p_yj),
  ncol=1,
  heights=c(1.0, 1.0),
  byrow=TRUE,
  draw=FALSE)
p
```

```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}

if(!file.exists(file.path(manuscript_dir, "residual_plot.RDS"))) stop("Residual data for goodness-of-fit test does not exist.")

data <- readRDS(file.path(manuscript_dir, "residual_plot.RDS"))

central_width <- c(0.60, 0.70, 0.80, 0.90, 0.95, 1.00)

p_lower <- 0.50 - central_width / 2
p_upper <- 0.50 + central_width / 2

x <- data.table::copy(data)

# Clip empirical probabilities to centre.
x <- x[p >= p_lower & p <= p_upper]

# Compute mean absolute residual error per feature.
x <- x[, list("mare"=mean(residual)), by=c("distribution_id", "method")]
x <- x[, list("n"=.N), by=c("mare", "method")][order(mare, method)]
x[, "rejected":= 1.0 - cumsum(n) / sum(n)]
x <- rbind(data.table::data.table("mare"=c(0.0, 0.0), "method"=c("box_cox", "yeo_johnson"), "n"=0L, "rejected"=0.0))

p <- ggplot2::ggplot(
  data = x,
  mapping = ggplot2::aes(
    x = mare,
    y = rejected,
    colour = method))
p <- p + plot_theme
p <- p + ggplot2::geom_step()
p <- p + ggplot2::xlab("mean absolute residual error")
p <- p + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE)
```

>>>>

```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Residual error as a function of empirical probability. ",
  "Percentiles of the error are shown for robust, shift-sensitive transformations of 1000 randomly drawn asymmetric ",
  "generalised normal distributions, with 10 randomly drawn outliers distributions, ",
  "(up to 10% of samples) each. Larger errors occur for more extreme values at the edges of each distribution. ",
  "For comparison, the 99th percentile of residual errors for transformations of outlier-free distributions are shown (grey dashed line).")

.plot_residuals(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir)

```{r goodness-of-fit, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Check")
.plot_
